A user-extensible and adaptable parser architecture

### 2008-08-14 doc/sgai-2008.tex
First batch of corrections from proofreading.

% XXX REMOVE THESE AND CHECK RESULTING LENGTH
\setlength{\floatsep}{0pt}
\setlength{\intextsep}{0pt}
\setlength{\textfloatsep}{0pt}

@@

The architecture described herein was developed as part of a larger project
to improve anti-spam techniques by analysing the performance of the set of
filters currently in use, optimising the order and membership of the set
based on that analysis, and developing supplemental filters where there is
a deficiency.  Most anti-spam techniques are
content-based~\cite{a-plan-for-spam, word-stemming, relaxed-online-svms},
which requires accepting the mail in the first place, but rejecting is
preferable in many cases: senders of non-spam mail that is mistakenly
rejected will receive an immediate non-delivery notice; load is reduced on
mail servers (allowing more intensive content-based techniques to be used);
users have less mail to wade through.  Improving the performance of
anti-spam techniques which are applied when mail is being transferred via
SMTP\footnotemark{} is the goal of this project, by providing a platform
for reasoning about anti-spam filters.  The approach chosen to measure
performance was to analyse the log files produced by the mail server,
rather than modifying the mail server to generate performance statistics:
this approach improves the chances of other sites testing and utilising the
software.  The need arose for a parser capable of dealing with the great
number and variety of log lines produced by Postfix~\cite{postfix}, the
SMTP server in use.  The parser must be designed so that adding support for
parsing new log lines is a simple task, as the log lines to be parsed will
change over time.  The variety in log lines occurs for several
reasons:\footnotetext{Simple Mail Transfer Protocol transfers mail from the
sender to one or more recipients across the Internet.  It is a simple,
human readable, plain text protocol, making it quite easy to test and debug
problems with it.  The original protocol definition is
RFC~821~\cite{RFC821}, updated in RFC~2821~\cite{RFC2821}.}

@@

It was hoped to reuse an existing parser rather than starting from scratch:
several parsers were considered and rejected because their parsing coverage
was too low, their parsing too inexact, and/or they did not extract
sufficient data.  The effort required to adapt and improve an existing
parser was judged to be greater than the effort to write one, as the
techniques used by the existing parsers severely limited their potential:
some ignored the majority of log lines, parsing specific log lines with
great accuracy, but without any provision for readily parsing new or
similar log lines; others sloppily parsed the majority of log lines, but
were incapable of distinguishing between log lines of the same type, e.g.\
rejections, and extracted only a portion of the data required.  The only
prior published work on the subject of parsing Postfix log files
is~\cite{log-mail-analyser}, which aims to present correlated log data in a
form suitable for a systems administrator to search using the myriad of
standard Unix text processing utilities already available.  A full state of
the art review is outside the scope of this paper but will be included in
the thesis resulting from this work.

The solution developed is conceptually simple: provide a few generic
functions (\textit{actions\/}), each capable of dealing with an entire
category of log lines (e.g.\ rejections), accompanied by a multitude of
precise patterns (\textit{rules\/}), each of which matches all log lines of
a specific type, and only that type (e.g.\ rejection by a specific DNSBL).
It is an accepted standard to separate the parsing procedure from the
declarative grammar it operates with.  Part of the novelty here is that the
grammars (rules and actions) are themselves procedures.  This architecture
is ideally suited to parsing inputs where the input is not fully understood
or does not follow a traditional grammar: the architecture warns about
unparsed inputs and other errors, but continues on parsing as best it can,
allowing the developer of a new parser to decide which deficiencies are
most important and require attention first, rather than being forced to fix
the first error which arises.

@@

    \item [Framework]  The framework is the structure which actions and
        rules fit into.  It provides the parsing loop, loading and
        validation of rules, shared data storage, storage of results, and
        other support functions.

@@

The framework tries each rule in turn until it finds one which matches the
input log line, then invokes the action specified by that rule.

@@

Decoupling also creates a clear separation of functionality: rules handle
low level details of identifying and extracting data from log lines;
actions handle the higher level details of assembling the required data,
dealing with the intricacies of the input being parsed, complications
arising, etc.;\ the framework provides services to actions and manages the
parsing process.

There is some similarity between the parser's design and William Wood's
Augmented Transition Networks (ATN)~\cite{atns, nlpip}, a tool used in
Computational Linguistics for creating grammars to parse or generate
sentences.  The resemblance between ATN and the parser is accidental, but
it is interesting how two different approaches have a similar division of
responsibilities, while having completely different implementations and
semantics.

@@

The framework takes care of miscellaneous support functions and low level
details of parsing, freeing the authors of actions to concentrate on
writing productive code.  It provides the link between actions and rules,
allowing either to be improved independently of the other.  It provides
shared storage to pass data between actions, loads and validates rules,
controls parsing of log files, invokes actions, tracks how often each rule
matches to optimise rule ordering (\sectionref{Rule ordering}), and stores
results in the database.  The framework is largely independent of the
parser being written: the same functionality will be required for most or
all parsers.  The framework is the core of the architecture and is
deliberately quite simple: the rules deal with the variation in log lines,
and the actions deal with the intricacies and complications encountered
when parsing.

@@

Each action contains the code required to deal with log lines from a
particular category, e.g.\ rejections.  The actions are parser-specific:
each parser author will need to write the actions required from scratch,
unless extending an existing parser.  In the Postfix parser developed for
this project there are \numberOFactions{} actions and \numberOFrules{}
rules; the distribution of rules to actions is very skewed, as shown in
\graphref{Distribution of rules per action}.  The distribution is skewed
for the same reason the architecture is so successful in enabling users to
parse new log lines: most new log lines can be parsed by a combination of a
new rule and an existing action.  It is expected that parsers utilising
this architecture will have a high ratio of rules to actions, hopefully
resulting in simpler rules and clearer distinctions between log lines.
\showgraph{build/graph-action-distribution}{Distribution
of rules per action}{Distribution of rules per action} Unsurprisingly the
action with the highest number of associated rules is \texttt{REJECTION},
the action which handles Postfix rejecting a mail; it is followed by
\texttt{SAVE\_BY\_QUEUEID}, the action responsible for handling informative
inputs: these are log lines that supplement the data gathered from other
log lines (such as rejections), but that do not by themselves represent
interesting events.  The third most common action is, perhaps surprisingly,
\texttt{IGNORE}: the \texttt{IGNORE} action does nothing when executed,
allowing uninteresting log lines to be parsed without causing any effects
(neither \texttt{REJECTION} nor \texttt{IGNORE} imply that the input is
ungrammatical or unparsed).  The parser warns about every unparsed log
line, to alert the user that they need to alter or extend their ruleset;
the Postfix parser successfully parses every input line in the
\numberOFlogFILESall{} log files it is currently being tested with.  The
remaining actions have only one or two associated rules: in some cases this
is because there will only ever be one type of log line for that action,
e.g.\ all log lines showing that a remote client has connected are matched
by a single rule and handled by the \texttt{CONNECT} action; in other cases
the action is required to address a deficiency in the log files, or a
complication which arises during parsing.

\label{example action}

Using the \texttt{CONNECT} action as an example, it creates a new data
structure for each client connecting, and saves the data extracted by the
rule.  If the data structure already exists it is treated as a symptom of a
bug, and the action issues a warning containing the full contents of the
data structure, plus the log line which has just been parsed.

@@

During development of the parser it became apparent that in addition to the
obvious variety in log lines, there were a large number of complications to
be overcome.  Some were the result of deficiencies in Postfix's logging
(some of which were improved on or resolved by later versions of Postfix);
others were due to the vagaries of process scheduling, client behaviour,
and administrative actions.  All were successfully accommodated in the
parser: adding new actions was enough for several of the complications;
others required modifications to a single existing action to work around
the difficulties; the remainder were resolved by adapting actions to pass
information to subsequent actions (via the framework), which check for the
extra information and change their behaviour appropriately.

Actions may also return a modified input line which will be parsed as if
read from the input stream, allowing for a simplified version of cascaded
parsing~\cite{cascaded-parsing}.  This is a powerful facility allowing a
generic rule and action to parse part of the input, followed by more
specific rules and actions parsing the remainder.

@@

Rules categorise inputs, specifying both the regex to match against each
log line and the action to invoke when the match is successful.  The
Postfix parser stores its rules in the same SQL database the results are
stored in, removing any doubt as to which set of rules was used to produce
a set of results; other implementations are free to store the rules in
whatever fashion suits their needs.

The framework requires that each rule have \texttt{action} and
\texttt{regex} attributes; each implementation is free to add any
additional attributes it requires.  The Postfix parser adds several
attributes to optimise rule ordering (\sectionref{Rule ordering}), to
restrict which log lines rules can be matched against, and to describe each
rule.  Data is automatically extracted and saved from the log line based on
keywords in the rule's regex, but there is also a facility to specify extra
data to be saved.

@@

The parser does not try to detect overlapping rules; that responsibility is
left to the author of the rules.  Unintentionally overlapping rules lead to
inconsistent parsing and data extraction because the first matching rule
wins, and the order in which rules are tried against each line may change
between log files.  Overlapping rules are frequently a requirement,
allowing a more specific rule to match some log lines and a more general
rule to match the majority, e.g.\ separating SMTP delivery to specific
sites from SMTP delivery to the rest of the world.  Overlapping rules
should have a priority attribute to specify the order in which they should
be used; negative priorities may be useful for catchall rules.

Storing the rules separately to the parser allows external tools to be
written to detect overlapping rules.  Traditional regexes are equivalent in
computational power to Finite Automata (FA); it is possible to convert
between regexes and FA, so regex overlap could be detected by finding a
non-empty intersection of two FA\@.  Perl 5.10 regexes are more powerful
than traditional regexes, e.g.\ it is possible to match correctly nested
and balanced brackets to an arbitrary depth:
\newline{}\tab{}\verb!/^[^<>]*(<(?:(?>[^<>]+)|(?1))*>)[^<>]*$/!\newline{}
Perl 5.10 regexes can maintain an arbitrary state stack and are thus
equivalent in computational power to Pushdown Automata (PA), so overlap
detection may require calculating the intersection of two PA.\@  Detecting
overlap between $n$ regexes requires $\frac{n(n-1)}{2}$ intersections to be
calculated, giving $O(n^2)$ complexity.  This is certainly not a task to be
performed on every parsing run: detecting overlap amongst the Postfix
parser's \numberOFrules{} rules would require \numberOFruleINTERSECTIONS{}
intersections to be calculated.

It is possible to define pathological regexes, which fall into two main
categories: regexes which match every input line, and regexes which consume
large amounts of CPU time during matching.  Defining a regex which matches
every input line is trivial: \verb!/^/! will match the start of every input
line.  In general a regex will not consume a large amount of CPU time when
the match is successful; usually CPU time is consumed by a regex with a lot
of alteration and variable quantifiers failing to match.  This topic is
beyond the scope of this paper, see~\cite{mastering-regular-expressions}
for in-depth discussion.

@@

The example rule in \tableref{Example rule} matches the message logged by
Postfix when a remote client connects to deliver mail; it would match the
following sample log line:

\begin{verbatim}
    connect from client.example.com[1.2.3.4]
\end{verbatim}

% Do not reformat this!
\begin{table}[htbp]
    \caption{Example rule}
    \empty{}\label{Example rule}
    \begin{tabular}[]{ll}
        \tableline{}%
        Attribute                 & Value                                            \\
        \tableline{}%
        program                   & \daemon{smtpd}                                   \\
        regex                     & \verb!^connect from (__HOSTNAME__)\[(__IP__)\]$! \\
        result\_cols              &                                                  \\
        connection\_cols~\empty{} & client\_hostname = 1; client\_ip = 2             \\
        action                    & CONNECT                                          \\
        priority                  & 0                                                \\
        \tableline{}%
    \end{tabular}
\end{table}

\noindent{}The \texttt{CONNECT} action is described in \sectionref{example
action}.

@@

    \item [Context-free rules:]  Rules are context free: they do not take
        into account inputs which have previously been seen, nor do they
        perform any kind of lookahead to future inputs.  In context-free
        grammar terms the parser rules could be described as:
        $\text{\textless{}input\textgreater{}} \mapsto \text{rule-1} |
        \text{rule-2} | \text{rule-3} | \dots | \text{rule-n}$

@@

    \item [Closer to Natural Language Processing than using a fixed
        grammar:] Unlike traditional parsers such as those used when
        compiling a programming language, this architecture does not have a
        fixed grammar specification which inputs must adhere to.  The
        architecture is capable of dealing with overlapping inputs, out of
        order inputs, and ambiguous inputs where heuristics must be applied
        --- all have arisen and been successfully dealt with in the Postfix
        log parser.

    \item [Line oriented:]  The architecture is line oriented at present:
        there is no facility for rules to consume more input or push unused
        input back onto the input stream.  This was not a deliberate
        decision, rather a consequence of the line oriented nature of the
        log files; more flexible approaches could be pursued.

@@

Parsing efficiency is an obvious concern when the parser routinely needs to
parse very large log files.  The server which generated the log files used
in testing this parser accepts approximately 10,000 mails for 700 users a
day.  Median log file size is 50MB, containing 285,000 log lines --- large
scale mail servers would have much larger log files.  When generating the
timing data used in this section, \numberOFlogFILES{} log files (totaling
10.08 GB, \numberOFlogLINEShuman{} log lines) were each parsed 10 times and
the parsing times averaged.  Saving results to the database was disabled
for the test runs, as that dominates the run time of the program, and the
tests are aimed at measuring the speed of the parser rather than the speed
of the database and the disks the database is stored on.  The computer used
for test runs was a Dell Optiplex 745 described in \tableref{Computer used
to generate statistics}.

@@

Parsing all \numberOFlogFILES{} log files in one run took
\input{build/include-timing-run-duration}.  The computer was dedicated to
the task of gathering statistics from test runs, and was not used for any
other purpose while test runs were ongoing; any services not necessary for
running the tests were disabled.

@@

An important property of a parser is how parsing time scales relative to
input size: does it scale linearly, polynomially, or exponentially?
\Graphref{Parsing time, file size, and number of lines} shows the parsing
time in seconds, file size in MB, and tens of thousands of log lines for
each of the \numberOFlogFILES{} log files.  All three lines run roughly in
parallel, giving the impression that the algorithm scales linearly with
input size.  This impression is borne out by \graphref{parsing time vs file
size vs number lines factor} which plots both the ratio of file size vs
parsing time, and the ratio of number of log lines vs parsing time (higher
is better in both cases).  The ratios are quite tightly banded across the
graph, showing that the algorithm scales linearly.  The ratio increases
(i.e.\ improves) for log files 22 and 62--68 despite their larger than
average size (shown in \graphref{Parsing time, file size, and number of
lines}).  \showgraph{build/graph-input-size-vs-parsing-time}{Parsing time,
file size, and number of lines}{Parsing time, file size, and number of
lines} \showgraph{build/graph-input-size-vs-parsing-time-ratio}{Ratio of
file size and number of lines to parsing time}{parsing time vs file size vs
number lines factor}

@@

\Graphref{Parsing time relative to shuffled} shows the parsing times of
optimal and reversed orderings relative to shuffled ordering.  Overall this
optimisation provides a modest but worthwhile reduction in parsing time of
approximately 10\%.  \showgraph{build/graph-hits}{Hits per rule}{rule hits
graph} \showgraph{build/graph-optimal-and-reverse-vs-shuffle}{Parsing time
relative to shuffled ordering}{Parsing time relative to shuffled}

@@

How any architecture scales as the number of rules increases is important,
but it is particularly important in this architecture because it is
expected that there will be a large number of rules.  There are
\numberOFrules{} rules in the full Postfix ruleset, whereas the minimum
number of rules required to parse the \numberOFlogFILES{} log files is
\numberOFrulesMINIMUM{}, \numberOFrulesMINIMUMpercentage{} of the total.  A
second set of statistics was generated using the minimum ruleset and the
results compared to the statistics generated using the full ruleset: the
percentage parsing time increase when using the full ruleset instead of the
minimal ruleset for optimal, shuffled and reversed orderings is shown in
\graphref{Percentage parsing time increase of maximum ruleset over minimum
ruleset}.
\showgraph{build/graph-full-ruleset-vs-minimum-ruleset}{Percentage parsing
time increase of maximum ruleset over minimum ruleset}{Percentage parsing
time increase of maximum ruleset over minimum ruleset}

It is clear from \graphref{Percentage parsing time increase of maximum
ruleset over minimum ruleset} that the increased number of rules has a
noticeable performance impact with reverse ordering, and to a lesser extent
with shuffled ordering.  The optimal ordering (where the most frequently
matching rules are tried first) shows a mean increase of
\input{build/include-full-ruleset-vs-minimum-ruleset} in parsing time for a
\numberOFrulesMAXIMUMpercentage{} increase in the number of rules.  These
results show that the architecture scales very well as the number of rules
increases, and that sorting the rules is an important optimisation enabling
this.

@@

The framework issues warnings for any log lines which are not parsed; no
warnings are issued for unparsed log lines while parsing the
\numberOFlogFILES{} test log files, so it can be safely concluded that
there are zero false negatives.  False positives are harder to quantify:
manually verifying that the correct regex parsed each of the 60,721,709 log
lines is infeasible, making it impossible to quantify the false positive
rate; however a random sample of 6039 log lines (0.00994\% of the total)
was parsed and the results verified to ensure that the correct regex parsed
each log line.

Coverage of mails is much more difficult to determine accurately than
coverage of log lines.  The parser can dump its state tables in a human
readable form; examining these tables with reference to the log files is
the best way to detect mails which were not handled properly.  The parser
issues warnings when it detects any errors, some of which may alert the
user to a problem.  There should be few or no warnings when parsing, and
when finished parsing the state table should only contain entries for mails
which had yet to be delivered when the log files ended, or were accepted
before the log files began.  A sample of 6000 log lines was parsed with all
debugging options enabled, resulting in 167,448 lines of output.  All
167,448 lines were examined in conjunction with the input log file and a
dump of the resulting database, verifying that for each of the input lines
the parser used the correct rule and executed the correct action, which in
turn produced the correct result and inserted the correct data in the
database.  The log segment produced 4 warnings (about mails which had
started before the log file segment), 10 mails remaining in the state
tables, and 1625 connections correctly entered in the database.  Given the
evidence detailed above, the authors are confident that the false positive
rate when reconstructing a mail is zero.

>>>>

% XXX REMOVE THESE AND CHECK RESULTING LENGTH
%\setlength{\floatsep}{0pt}
%\setlength{\intextsep}{0pt}
%\setlength{\textfloatsep}{0pt}

@@

The architecture described herein was developed as part of a larger project
to improve anti-spam techniques by analysing the performance of the set of
filters currently in use, optimising the order and membership of the set
based on that analysis, and developing supplemental filters where there is
a deficiency.  Most anti-spam techniques are
content-based~\cite{a-plan-for-spam, word-stemming, relaxed-online-svms},
which requires accepting the mail in the first place, but rejecting mail
during the delivery attempt is preferable: senders of non-spam mail that is
mistakenly rejected will receive an immediate non-delivery notice; load is
reduced on mail servers (allowing more intensive content-based techniques
to be used); users have less mail to wade through.  Improving the
performance of anti-spam techniques which are applied when mail is being
transferred via SMTP\footnotemark{} is the goal of this project, by
providing a platform for reasoning about anti-spam filters.  The approach
chosen to measure performance was to analyse the log files produced by the
mail server, rather than modifying the mail server to generate performance
statistics: this approach improves the chances of other sites testing and
utilising the software.  The need arose for a parser capable of dealing
with the great number and variety of log lines produced by
Postfix~\cite{postfix}, the SMTP server in use.  The parser must be
designed so that adding support for parsing new log lines is a simple task,
as the log lines to be parsed will change over time.  The variety in log
lines occurs for several reasons:\footnotetext{Simple Mail Transfer
Protocol transfers mail from the sender to one or more recipients across
the Internet.  It is a simple, human readable, plain text protocol, making
it quite easy to test and debug problems with it.  The original protocol
definition is RFC~821~\cite{RFC821}, updated in RFC~2821~\cite{RFC2821}.}

@@

It was hoped to reuse an existing parser rather than starting from scratch:
several parsers were considered and rejected because their parsing coverage
was too low, their parsing too inexact, and/or they did not extract
sufficient data.  The effort required to adapt and improve an existing
parser was judged to be greater than the effort to write one, as the
techniques used by the existing parsers severely limited their potential:
some ignored the majority of log lines, parsing specific log lines with
great accuracy, but without any provision for readily parsing new or
similar log lines; others sloppily parsed the majority of log lines, but
were incapable of distinguishing between log lines of the same type, e.g.\
rejections, and extracted only a small portion of the data required.  The only
prior published work on the subject of parsing Postfix log files
is~\cite{log-mail-analyser}, which aims to present correlated log data in a
form suitable for a systems administrator to search using the myriad of
standard Unix text processing utilities already available.  A full state of
the art review is outside the scope of this paper but will be included in
the thesis resulting from this work.

The solution developed is conceptually simple: provide a few generic
functions (\textit{actions\/}), each capable of dealing with an entire
category of log lines (e.g.\ rejections), accompanied by a multitude of
precise patterns (\textit{rules\/}), each of which matches all log lines of
a specific type, and only that type (e.g.\ rejection by a specific DNSBL).
XXX IMPROVE:\@ It is an accepted standard to separate the parsing procedure
from the declarative grammar it operates with.  Part of the novelty here is
that the grammars (rules and actions) are themselves procedures.  This
architecture is ideally suited to parsing inputs where the input is not
fully understood or does not conform to a fixed grammar: the architecture
warns about unparsed inputs and other errors, but continues on parsing as
best it can, allowing the developer of a new parser to decide which
deficiencies are most important and require attention first, rather than
being forced to fix the first error which arises.

@@

    \item [Framework]  The framework is the structure which actions and
        rules plug into.  It provides the parsing loop, loading and
        validation of rules, shared data storage, storage of results, and
        other support functions.

@@

For each input log line the framework tries each rule in turn until it
finds one which matches the input log line, then invokes the action
specified by that rule.

@@

Decoupling also creates a clear separation of functionality: rules handle
low level details of identifying log lines and extracting data; actions
handle the higher level details of assembling the required data, dealing
with the intricacies of the input being parsed, complications arising,
etc.;\ the framework provides services to actions and manages the parsing
process.

There is some similarity between the architecture and William Wood's
Augmented Transition Networks (ATN)~\cite{atns, nlpip}, a tool used in
Computational Linguistics for creating grammars to parse or generate
sentences.  The resemblance between ATN and the parser is accidental, but
it is interesting how two different approaches have a similar division of
responsibilities, while having different semantics.

@@

The framework takes care of miscellaneous support functions and low level
details of parsing, freeing the authors of actions to concentrate on
writing productive code.  It provides the link between actions and rules,
allowing either to be improved independently of the other.  It provides
shared storage to pass data between actions, loads and validates rules,
controls parsing of log files, invokes actions, tracks how often each rule
matches to optimise rule ordering (\sectionref{Rule ordering}), and stores
results in the database.  The framework is largely independent of the
parser being written: the same functionality will be required for most or
all parsers, though some specialised support functions may be provided.
The framework is the core of the architecture and is deliberately quite
simple: the rules deal with the variation in log lines, and the actions
deal with the intricacies and complications encountered when parsing.

@@

Each action contains the code required to deal with a particular type of
log lines, e.g.\ rejections.  The actions are parser-specific: each parser
author will need to write the actions required from scratch, unless
extending an existing parser.  In the Postfix parser developed for this
project there are \numberOFactions{} actions and \numberOFrules{} rules,
with a very skewed distribution of rules to actions as shown in
\graphref{Distribution of rules per action}.  The distribution is skewed
for the same reason the architecture is so successful in enabling users to
parse new log lines: most new log lines can be parsed by a combination of a
new rule and an existing action.  It is expected that parsers utilising
this architecture will have a high ratio of rules to actions, hopefully
resulting in simpler rules and clearer distinctions between log lines.
\showgraph{build/graph-action-distribution}{Distribution of rules per
action}{Distribution of rules per action} Unsurprisingly the action with
the highest number of associated rules is \texttt{REJECTION}, the action
which handles Postfix rejecting a mail; it is followed by
\texttt{SAVE\_BY\_QUEUEID}, the action responsible for handling informative
inputs: these are log lines that supplement the data gathered from other
log lines (such as rejections), but that do not by themselves represent
interesting events.  The third most common action is, perhaps surprisingly,
\texttt{IGNORE}: the \texttt{IGNORE} action does nothing when executed,
allowing uninteresting log lines to be parsed without causing any effects
(neither \texttt{REJECTION} nor \texttt{IGNORE} imply that the input is
ungrammatical or unparsed).  The parser warns about every unparsed log
line, to alert the user that they need to alter or extend their ruleset;
the Postfix parser successfully parses every input line in the
\numberOFlogFILESall{} log files it is currently being tested with.  The
remaining actions have only one or two associated rules: in some cases this
is because there will only ever be one type of log line for that action,
e.g.\ all log lines showing that a remote client has connected are matched
by a single rule and handled by the \texttt{CONNECT} action; in other cases
the action is required to address a deficiency in the log files, or a
complication which arises during parsing.  Using the \texttt{CONNECT}
action as an example, it creates a new data structure for each client
connecting, and saves the data extracted by the rule.  If the data
structure already exists it is treated as a symptom of a bug, and the
action issues a warning containing the full contents of the data structure,
plus the log line which has just been parsed.

@@

During development of the Postfix parser it became apparent that in
addition to the obvious variety in log lines, there were a large number of
complications to be overcome.  Some were the result of deficiencies in
Postfix's logging (some of which were improved on or resolved by later
versions of Postfix); others were due to the vagaries of process
scheduling, client behaviour, and administrative actions.  All were
successfully accommodated in the Postfix parser: adding new actions was
enough to overcome several of the complications; others required
modifications to a single existing action to work around the difficulties;
the remainder were resolved by adapting actions to pass information to
subsequent actions (via the framework), which check for the extra
information and change their behaviour appropriately.

Actions may also return a modified input line which will be parsed as if
read from the input stream, allowing for a simplified version of cascaded
parsing~\cite{cascaded-parsing}.  This is a powerful facility allowing a
number of rules and actions to parse a single input line.

@@

Rules categorise inputs, specifying both the regex to match against each
log line and the action to invoke when the match is successful.  The
Postfix parser stores its rules in the same SQL database the results are
stored in, removing any doubt as to which set of rules was used to produce
a set of results; other implementations are free to store the rules in
whatever fashion suits their needs.  The framework requires each rule to
have \texttt{action} and \texttt{regex} attributes; each implementation is
free to add any additional attributes it requires.  The Postfix parser adds
several attributes to optimise rule ordering (\sectionref{Rule ordering}),
to restrict which log lines rules can be matched against, and to describe
each rule.  Data is automatically extracted and saved from the log line
based on keywords in the rule's regex, but there is also a facility to
specify extra data to be saved.

@@

The parser does not try to detect overlapping rules: that responsibility is
left to the author of the rules.  Unintentionally overlapping rules lead to
inconsistent parsing and data extraction because the first matching rule
wins, and the order in which rules are tried against each line may change
between log files.  Overlapping rules are frequently a requirement,
allowing a more specific rule to match some log lines and a more general
rule to match the majority, e.g.\ separating SMTP delivery to specific
sites from SMTP delivery to the rest of the world.  Overlapping rules
should have a priority attribute to specify the order in which they should
be used; negative priorities may be useful for catchall rules.

Storing the rules separately to the parser allows external tools to be
written to detect overlapping rules.  Traditional regexes are equivalent in
computational power to Finite Automata (FA); it is possible to convert
between regexes and FA, so regex overlap could be detected by finding a
non-empty intersection of two FA\@.  Perl 5.10 regexes are more powerful
than traditional regexes, e.g.\ it is possible to match correctly nested
and balanced brackets to an arbitrary depth:
\newline{}\tab{}\verb!/^[^<>]*(<(?:(?>[^<>]+)|(?1))*>)[^<>]*$/!\newline{}
Perl 5.10 regexes can maintain an arbitrary state stack and are thus
equivalent in computational power to Pushdown Automata (PA), so overlap
detection may require calculating the intersection of two PA.\@  Detecting
overlap between $n$ regexes requires $\frac{n(n-1)}{2}$ intersections to be
calculated, giving $O(n^2x)$ complexity, where $O(x)$ is the complexity of
calculating intersection.  This is certainly not a task to be performed on
every parsing run: detecting overlap amongst the Postfix parser's
\numberOFrules{} rules would require calculating
\numberOFruleINTERSECTIONS{} intersections.

It is possible to define pathological regexes, which fall into two main
categories: regexes which match every input line, and regexes which consume
large amounts of CPU time during matching.  Defining a regex which matches
every input line is trivial: \verb!/^/! matches the start of every input
line.  In general a regex will not consume a large amount of CPU time when
the match is successful; usually CPU time is consumed by a regex with a lot
of alteration and variable quantifiers failing to match.  This topic is
beyond the scope of this paper, see~\cite{mastering-regular-expressions}
for in-depth discussion.

@@

The example rule in \tableref{Example rule} matches the following sample
log line logged by Postfix when a remote client connects to deliver mail:
\newline{}\tab{}\verb!connect from client.example.com[1.2.3.4]!

% Do not reformat this!
\begin{table}[htbp]
    \caption{Example rule}
    \empty{}\label{Example rule}
    \begin{tabular}[]{ll}
        \tableline{}%
        Attribute                 & Value                                            \\
        \tableline{}%
        %program                   & \daemon{smtpd}                                   \\
        regex                     & \verb!^connect from (__HOSTNAME__)\[(__IP__)\]$! \\
        action                    & CONNECT                                          \\
        %priority                  & 0                                                \\
        \tableline{}%
    \end{tabular}
\end{table}

@@

    \item [Context-free rules:]  Rules are context free: they do not take
        into account past or future inputs.  In context-free
        grammar terms the parser rules could be described as:
        $\text{\textless{}input\textgreater{}} \mapsto \text{rule-1} |
        \text{rule-2} | \text{rule-3} | \dots | \text{rule-n}$

@@

    \item [Closer to Natural Language Processing than using a fixed
        grammar:] Unlike traditional parsers such as those used when
        compiling a programming language, this architecture does not have a
        fixed grammar specification which inputs must adhere to.  The
        architecture is capable of dealing with interleaved inputs, out of
        order inputs, and ambiguous inputs where heuristics must be applied
        --- all have arisen and been successfully dealt with in the Postfix
        log parser.

    \item [Line oriented:]  The architecture is line oriented at present:
        there is no facility for rules to consume more input or push unused
        input back onto the input stream (however actions may change input
        lines and cause the modified line to be re-parsed).  This was not a
        deliberate decision, rather a consequence of the line oriented
        nature of the log files; more flexible approaches could be pursued.

@@

Parsing efficiency is an obvious concern when the parser routinely needs to
parse very large log files.  The mail server which generated the log files
used in testing this parser accepts approximately 10,000 mails for 700
users a day.  Median log file size is 50MB, containing 285,000 log lines
--- large scale mail servers would have much larger log files.  When
generating the timing data used in this section, \numberOFlogFILES{} log
files (totaling 10.08 GB, \numberOFlogLINEShuman{} log lines) were each
parsed 10 times and the parsing times averaged.  Saving results to the
database was disabled for the test runs, as that dominates the run time of
the program, and the tests are aimed at measuring the speed of the parser
rather than the speed of the database and the disks the database is stored
on.  The computer used for test runs was a Dell Optiplex 745 described in
\tableref{Computer used to generate statistics}.  Parsing all
\numberOFlogFILES{} log files in one run took
\input{build/include-timing-run-duration}.  The computer was dedicated to
the task of gathering statistics from test runs, and was not used for any
other purpose while test runs were ongoing; any services not necessary for
running the tests were disabled.

@@

@@

An important property of a parser is how parsing time scales relative to
input size: does it scale linearly, polynomially, or exponentially?
\Graphref{Parsing time, file size, and number of lines} shows the parsing
time in seconds, file size in MB, and number of log lines in tens of
thousands, for each of the \numberOFlogFILES{} log files.  All three lines
run roughly in parallel, giving the impression that the algorithm scales
linearly with input size.  This impression is borne out by
\graphref{parsing time vs file size vs number lines factor} which plots
both the ratio of file size vs parsing time, and the ratio of number of log
lines vs parsing time (higher is better in both cases).  The ratios are
quite tightly banded across the graph, showing that the algorithm scales
linearly.  The ratio increases (i.e.\ improves) for log files 22 and 62--68
despite their larger than average size (shown in \graphref{Parsing time,
file size, and number of lines}).

\showgraph{build/graph-input-size-vs-parsing-time}{Parsing time, file size,
and number of lines}{Parsing time, file size, and number of lines}
\showgraph{build/graph-input-size-vs-parsing-time-ratio}{Ratio of number of
lines and file size to parsing time}{parsing time vs file size vs number
lines factor}

@@

\Graphref{Parsing time relative to shuffled} shows the parsing times of
optimal and reversed orderings as a percentage of the parsing time of
shuffled ordering.  Overall this optimisation provides a modest but
worthwhile reduction in parsing time of approximately 10\%.
\showgraph{build/graph-hits}{Hits per rule}{rule hits graph}
\showgraph{build/graph-optimal-and-reverse-vs-shuffle}{Parsing time
relative to shuffled ordering}{Parsing time relative to shuffled}

@@

How any architecture scales as the number of rules increases is important,
but it is particularly important in this architecture because it is
expected that there will be a large number of rules.  There are
\numberOFrules{} rules in the full Postfix parser ruleset, whereas the
minimum number of rules required to parse the \numberOFlogFILES{} log files
is \numberOFrulesMINIMUM{}, \numberOFrulesMINIMUMpercentage{} of the full
ruleset.  A second set of statistics was generated using the minimum
ruleset and the parsing times compared to those generated using the full
ruleset: the percentage parsing time increase when using the full ruleset
instead of the minimal ruleset for optimal, shuffled and reversed orderings
is shown in \graphref{Percentage parsing time increase of maximum ruleset
over minimum ruleset}.
\showgraph{build/graph-full-ruleset-vs-minimum-ruleset}{Percentage parsing
time increase of maximum ruleset over minimum ruleset}{Percentage parsing
time increase of maximum ruleset over minimum ruleset}

It is clear from \graphref{Percentage parsing time increase of maximum
ruleset over minimum ruleset} that the increased number of rules has a
noticeable performance impact with reverse ordering, and a lesser impact
with shuffled ordering.  The optimal ordering shows a mean increase of
\input{build/include-full-ruleset-vs-minimum-ruleset} in parsing time for a
\numberOFrulesMAXIMUMpercentage{} increase in the number of rules.  These
results show that the architecture scales very well as the number of rules
increases, and that sorting the rules is an important optimisation enabling
this.

@@

Warnings are issued for any log lines which are not parsed; no
warnings are issued for unparsed log lines while parsing the
\numberOFlogFILES{} test log files, so it can be safely concluded that
there are zero false negatives.  False positives are harder to quantify:
manually verifying that the correct regex parsed each of the
\numberOFlogLINES{} log lines is infeasible, making it impossible to
quantify the false positive rate; however a random sample of 6039 log lines
(0.00994\% of the total) was parsed and the results manually verified by
inspection to ensure that the correct regex parsed each log line.

Coverage of mails is much more difficult to determine accurately than
coverage of log lines.  The parser can dump its state tables in a human
readable form; examining these tables with reference to the log files is
the best way to detect mails which were not handled properly.  The parser
issues warnings when it detects any errors, some of which may alert the
user to a problem.  There should be few or no warnings when parsing, and
when finished parsing the state table should only contain entries for mails
which had yet to be delivered when the log files ended, or were accepted
before the log files began.  A sample of 6000 log lines was parsed with all
debugging options enabled, resulting in 167,448 lines of output.  All
167,448 lines were examined in conjunction with the input log file and a
dump of the resulting database, verifying that for each of the input lines
the parser used the correct rule and executed the correct action, which in
turn produced the correct result and inserted the correct data in the
database.  The log segment produced 4 warnings (about mails which had
started before the log file segment), 10 mails remaining in the state
tables, and 1625 connections correctly entered in the database.  No error
messages were produced, therefore there were no false negatives.  Given the
evidence detailed above, the authors are confident that the false positive
rate when reconstructing a mail is zero.
